Bayesian Shape Reconstruction and Optimal Guidance for Autonomous Landing on Asteroids

### 2021-12-19 manuscript.tex
Update manuscript with comments

\begin{abstract}
    Constructing of the precise shape of an asteroid is critical for spacecraft operations, as the gravitational potential is determined by spatial mass distribution.
    %For example, the standard method of determining the gravitational potential of an asteroid, through the use of a polyhedron potential model, is directly dependent on the shape model.
    %Furthermore, safe landing or close-proximity operations requires accurate knowledge of the surface topology. 
    The typical approach to shape determination requires a prolonged ``mapping'' phase of the mission over which extensive measurements are collected and transmitted for Earth-based processing.
    This paper presents a set of approaches to explore an unknown asteroid with onboard calculations, and to land on its surface area selected in an optimal fashion. 
    The main motivation is to avoid the extended period of mapping phase or preliminary ground observations that are commonly required in spacecraft mission around asteroids. 
    First, range measurements from spacecraft to the surface are used to incrementally correct an initial shape estimate according to the Bayesian framework. 
    Then, an optimal guidance scheme is proposed to control the vantage point of the range sensor to construct a complete 3D model of the asteroid shape. 
    This shape model is then used in a nonlinear controller to track a desired trajectory about the asteroid.
    Finally, a multi resolution approach is presented to construct a higher fidelity shape representation in a specified location while avoiding the inherent burdens of a uniformly high resolution mesh. 
    This approach enables for an accurate shape determination around a potential landing site.
    We demonstrate this approach using several radar shape models of asteroids and provide a full dynamic simulation about asteroid 4769 Castalia.
    \keywords{Asteroids \and Shape Construction \and Guidance \and Optimization}
\end{abstract}

@@

% dynamics are difficult around asteroids
The dynamic environment around asteroids is strongly perturbed and challenging for analysis and mission operations~\cite{scheeres2012}.
Due to their low mass, which in turn causes a low gravitational attraction, asteroids may have irregular shapes.
Furthermore, asteroids may also have a chaotic spin state due to the absorption and emittance of solar radiation~\cite{rubincam2000}.
As a result, approaches utilizing an inverse square gravitational model do not capture the  true dynamic environment.
In addition, the vast majority of asteroids are difficult to track and characterize using ground based sensors.
Due to their small size, frequently with a maximum radius less than \SI{1}{\kilo\meter}, and low albedo, the reflected energy of these asteroids is insufficient for reliable detection or tracking.
Therefore, the dynamics model of the asteroid is relatively coarse prior to in situ measurements from a dedicated spacecraft.
As a result, any spacecraft mission to an asteroid must include the ability to update the dynamic model given in situ measurements and remain robust to unmodelled forces.

Another key dynamic consideration is the coupling between rotational and translational states around the asteroid.
The coupling is induced due to the different gravitational forces experienced on various portions of the spacecraft. 
The effect of the gravitational coupling is related to the ratio of the spacecraft size and orbital radius~\cite{hughes2004}.
For operations around asteroids, the ratio is relatively large which causes a much larger coupling between the translational and rotational states.
In~\cite{elmasri2005} and \cite{sanyal2004a}, the coupling of an elastic dumbbell spacecraft in orbit about a central body is considered, assuming that the central body is spherically symmetric.
Furthermore, the spacecraft model is assumed to remain in a planar orbit.
As a result, these developments are not directly applicable to motion about an asteroid, which experiences highly non-Keplerian motion.
In~\cite{misra2015b}, the effect of coupled motion is investigated for long term trajectories around asteroids.
However, the analysis only considered a second order spherical harmonic gravitational potential model. 
Therefore, these results are only valid when far from the asteroid surface and will diverge when used within the Brillouin sphere.

@@

% on arrival spacecraft spend long periods mapping, and depending on mission this might be unallocable
On approach to an asteroid, spacecraft navigation and guidance is primarily based on ground measurements.
After arrival, a spacecraft will generally spend months or years in a mapping mission phase~\cite{kubota2003,cole1998}.
During this period, spacecraft sensors, such as on board optical telescopes or Light radio Detection and Ranging (LIDAR), are used to characterize the asteroid.
The resulting imagery and range data is transmitted to the ground and the resulting asteroid shape and motion is estimated. 
During this mapping phase the spacecraft must remain in a quiescent state devoted entirely to mapping the surface.
Depending on the mission type, this long period of mapping is crucial to the mission, such as sample collection~\cite{gates2015}. 
However, other missions, such as asteroid mitigation, may be severely limited by the time and ground resources required to generate a surface shape.
Furthermore, the long distances involved necessitate on-board autonomy to enable to spacecraft to operate without ground communications.
Similarly, during landing the spacecraft will require the ability to sense and model the surface topography in order to safely land in an unknown environment.
The dependency on expensive ground-based shape reconstruction techniques limits the ability of spacecraft to autonomously operate at asteroids.
Mitigation of this ground based surface modeling will greatly expand the range of missions possible.

% TODO Reduce some of these paragraphs since they're wordy
% we want to generate the shape in real time and then use this shape for updated and better control
This paper presents three techniques to explore an asteroid without need for an extended period of mapping phase or ground based measurements. 
First, we develop a method to compute the surface shape of an asteroid from onborad range measurements.
Assuming that a meshed model of an initial estimate, such as an ellipsoidal shape, is available, the radius of any vertex in the measured area is adjusted according to the measurements. 
This is completed in a stochastic fashion where the degree of confidence in the current shape is compared with the level of measurement noise. 
Our approach is able to operate in real time and incrementally update the shape model of an asteroid as new range measurements are collected.
This approach allows for the shape to be continually updated as range measurements are used to locally modify the shape estimate.
This technique is verified by numerical examples of shape construction for asteroids Geographos and Golevka.

@@

The optimal guidance and multi resolution mapping schemes are illustrated by dynamic simulation of landing on the asteroid Castalia. 
During this simulation, This updated shape model is then used to in a nonlinear controller to track a desired state trajectory for the dynamics of a rigid body spacecraft.
The dynamics are developed on the nonlinear manifold of rigid body motions, namely the special Euclidean group.
This formulation is based on an intrinsic geometric description of the motion and accurately captures the coupling between orbit and attitude dynamics.

@@

In this paper, we consider the motion of a dumbbell model of spacecraft around an asteroid.
The dumbbell model captures the important interactions of the coupling between orbital and attitude dynamics.
The dumbbell is defined by two spherical masses of radius \( r_1, r_2 \in \R \) with masses \( m_1, m_2 \in  \R\).
The masses are separated by a massless rod of length \( l \) and attached to the centers of each mass.
\Cref{fig:dumbbell_sc} shows the model and associated parameters.
\begin{figure}[htbp]
    \centering
    \input{dumbbell.tikz}
    \caption{Dumbbell Model of Rigid Spacecraft\label{fig:dumbbell_sc}}
\end{figure}
Without loss of generality, we define body fixed frames for both the spacecraft and asteroid, which are aligned with the principle axes of each body and originate at their respective center of mass. 
The spacecraft body fixed frame is centered at the center of mass of the vehicle.
The \( b_1 \)  axis is aligned with the connecting rod and directed along the axis of symmetry.
The \( b_2, b_3 \) axes span the plane orthogonal to the axis of symmetry of the dumbbell.
The distance from the center of mass to each spherical mass is defined as
\begin{align}\label{eq:dumbbell_mass_distances}
    l_1 &= \frac{m_2}{m_1 + m_2} l, \\
    l_2 &= l - l_1.
\end{align}
The asteroid is modeled as a constant density polyhedron with constant, and known, spin about the axis of its maximum moment of inertia. 
The axes of the body fixed frame for the asteroid are denoted by $\{f_1,f_2,f_3\}$. 
We also define the inertial frame, whose axes are denoted by $\{e_1,e_2,e_3\}$.

@@

\subsection{Spacecraft Dynamic Model}

@@

Here, we summarize the key equations required for implementation.
The polyhedron potential is defined as 
\begin{align}
    U(\vc{r}) &= \frac{1}{2} G \sigma \sum_{e \in \text{edges}} \vc{r}_e \cdot \vc{E}_e \cdot \vc{r}_e \cdot L_e - \frac{1}{2}G \sigma \sum_{f \in \text{faces}} \vc{r}_f \cdot \vc{F}_f \cdot \vc{r}_f \cdot \omega_f,
\end{align}
and \( \vc{r}_e\) and \(\vc{r}_f \) are the vectors from the spacecraft to any point on the respective edge or face, \( G\in\Re\) is the universal gravitational constant, and \( \sigma\in\Re \) is the constant density of the asteroid.
We define the attraction, gravity gradient matrix, and Laplacian as
\begin{align}
    \nabla U ( \vc{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vc{E}_e \cdot \vc{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \vc{r}_f \cdot \omega_f \in \R^{3 \times 1} , \label{eq:attraction}\\
    \nabla \nabla U ( \vc{r} ) &= G \sigma \sum_{e \in \text{edges}} \vc{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \omega_f \in \R^{3 \times 3}, \label{eq:gradient_matrix}\\
    \nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f \in \R^1 .\label{eq:laplacian}
\end{align}
The Laplacian, defined in~\cref{eq:laplacian}, gives a simple method to determine if the spacecraft has collided with the body~\cite{werner1996}. 
These equations are utilized to compute the gravitational potential of  the presented dumbbell spacecraft model.

@@

We assume the spacecraft contains a range sensor, such as LIDAR, that allows for the accurate measurement of the relative distance between the spacecraft and asteroid~\cite{zuber1997,zuber2000}.
This type of sensor measures the round-trip time for a pulse of energy to leave the spacecraft, reflect off the surface, and return to a collector on board.
Given the time total time of flight, the distance can be accurately computed using \( d = \frac{\Delta TOF}{2 c} \) where \( c = \SI{2.998e8}{\meter\per\second}\) is the constant speed of light.
Assuming accurate knowledge of the pointing direction of the spacecraft, in the form of the rotation matrix \( R \in \R^{3 \times 3 } \), we can compute a direction from the spacecraft to the measurement location on the surface.
The output of this sensor is a vector, \( \vc{d}_i\in\Re^3 \), defined in the spacecraft fixed frame which gives the direction to a measurement point on the surface. 
Using the state of the asteroid, we can transform this measurement to the asteroid fixed frame using the simple transformation
\begin{align*}
    \vc{p}_i = R_A^T (x+R \vc{d}_i), 
\end{align*}
which is the vector from the origin of the asteroid fixed frame to the surface point of the measurement.

@@

Our algorithm applies a probabilistic framework to radially modify each vertex \( \vc{v}_i \in \R^3\) of the shape estimate based on measurement \( \vc{p}_i \in \R^3 \). 
In other words, each vertex from the initial estimate of shape is either stretched or shrunk along its radial direction, without any rotation. 
This framework is reasonable assuming that both of the initial estimate and the actual shape does not have any hole. 
This approach alleviates much of complexity of incorporating new vertices or surface triangulation common in surface reconstruction methods~\cite{berg2008}.
This implies that the total number of vertices of the shape model is fixed.
However, additional detail, in the form of additional vertices, is possible by using standard mesh subdivision algorithms~\cite{orourke1998}, which is discussed in the subsequent section for multi resolution mapping.

@@

From spherical trigonometry~\cite{gade2010}, the central angle between measurement \( \vc{p}_j \) and vertex \( \vc{v}_i \) of the shape estimate is given by
\begin{align}\label{eq:geodesic_distance}
    \Delta \sigma_{j,i} = \arctan \parenth{\frac{\norm{\vc{p}_j \times \vc{v}_i}}{\vc{p}_j \cdot \vc{v}_i }}.
\end{align}

@@

The results in this section utilize a kinematics only model of the spacecraft and dynamics instead of the full dynamic simulation. 
We ignore the dynamics of the asteroid and spacecraft and instead focus solely on the shape reconstruction, and the spacecraft is assumed to be able to arbitrarily move around the asteroid and collect measurements.
The spacecraft remains at a fixed orbital radius defined as twice the maximum semi-major axis of the asteroid.
The simulations begin with an triaxial ellipsoid mesh that is sized to match the semi-major axes of each asteroid.
With this estimate, measurements are made of the surface and used to reconstruct the true shape following the process in~\cref{sec:radius_update}.
LIDAR measurements are generated until the total uncertainty of the model
\begin{align*}
    \sum_i w_i,
\end{align*}
is sufficiently small.
This uncertainty metric is the summation of the uncertainty of each vertex of the shape estimate.
As a result, a small total uncertainty is used to indicate that additional measurements would not substantially modify the shape.
In addition, we compute the volume of the estimated shape and compare against the volume of the truth shape model~\cite{neese2004}.

@@

The control gains are chosen based on the desired closed-loop system response. 
A variety of techniques are available to choose these gains, but a simple linear analysis offers a straightforward and systematic approach to choosing suitable values. 
We use the control inputs defined in~\cref{eq:translational_control,eqn:nodist_control} and substitute them into the dynamic equations of motion in~\cref{eq:x_dot,eq:v_dot,eq:R_dot,eq:omega_dot}.

@@

We utilize radar shape models of asteroids \num{4769} Castalia and (\num{52760}) \num{1998} \(\text{ML}_{14}\).
The examples demonstrate the full dynamic simulation of a rigid spacecraft with an autonomous closed loop control scheme to both reconstruct the asteroid shape.
The proposed optimal guidance scheme is executed in the outer-loop to generate the desired trajectory, which is followed by the nonlinear control scheme in the inner-loop. 
More specifically, the nonlinear controllers described previously are used to control both the translational and rotational states of the vehicle, where the control inputs are computed using the current shape estimate of the asteroid. 
As measurements are collected, the spacecraft autonomously updates its shape estimate and uses this estimate to compute the control inputs and desired future states.
As such, both the shape model and the gravitational potential available to the controller will be gradually refined. 
However, throughout the simulation, the actual spacecraft dynamics are computed with the gravity computed by the full shape model unknown to the controller.
These demonstrate the ability of a spacecraft to autonomously explore and maneuver around an initially-unknown asteroids while incrementally updating the shape model.

The asteroids are assumed to constantly rotate about the \( \vc{f}_3 = \vc{e}_3\) axis according to the parameters given in~\cref{tab:dynamic_asteroids}.
Furthermore, the state of the asteroid, namely the rotation matrix \( R_A \), is assumed to be known based on ground measurements or previous data.
\begin{table}[htbp]
    \centering
    \begin{tabular}{lcc}
        \toprule
        Property & \num{4769} Castalia & (\num{52760}) \num{1998} \(\text{ML}_{14}\) \\
        \midrule
        Semi-major axes(\si{\kilo\meter}) & \( 0.8065 \times 0.4905 \times 0.413 \) & \( 1.1 \times 1.1 \times 1.1 \) \\
        Rotational Period (\si{\hour}) & \num{4.095} & \num{14.98} \\
        Density (\si{\gram\per\centi\meter^3}) & \num{2.1} & \num{2.1} \\
        Vertices & \num{2048}  & \num{8192} \\
        Faces & \num{4092} & \num{16320} \\
        \bottomrule
    \end{tabular}
    \caption{Asteroid properties for dynamic exploration~\label{tab:dynamic_asteroids}}
\end{table}
At the beginning of the simulation the spacecraft is assumed to lie on the inertial \( e_1 \) axis, i.e.\ \( x_0 = [ x_0,\, 0,\,  0 ]\, \si{\kilo\meter} \).
In addition, at the initial state the spacecraft is orientated such that the \( b_1 \) axis is aligned with the inertial \( e_2 \) axis.
In other words the initial orientation is given by \( R_0 = \exp(\frac{\pi}{2} \hat{e}_3)\).
The shape reconstruction phase of the simulation is performed over \SI{15000}{\second}, over which time the spacecraft will take LIDAR measurements of the surface at \SI{1}{\hertz}.
Once the total uncertainty has been reduced sufficiently the spacecraft maneuvers to a ``home'' position aligned with the \( f_1 \) axis of the asteroid.

@@

Utilizing these metrics, a landing site is chosen to minimize the surface cost given as
\begin{align}\label{eq:surface_cost}
    \min_{v_i} J_l(x, v_i) =  J_{\text{distance}}(x, v_i) - J_{\text{science}}(v_i),
\end{align}
subject to a hard inequality constraint requiring that the surface slope is less than a threshold, i.e., \( \phi \leq \phi_m \).

>>>>

\begin{abstract}
    Constructing of the precise shape of an asteroid is critical for spacecraft operations, as the gravitational potential is determined by spatial mass distribution.
    %For example, the standard method of determining the gravitational potential of an asteroid, through the use of a polyhedron potential model, is directly dependent on the shape model.
    %Furthermore, safe landing or close-proximity operations requires accurate knowledge of the surface topology. 
    The typical approach to shape determination requires a prolonged ``mapping'' phase of the mission over which extensive measurements are collected and transmitted for Earth-based processing.
    This paper presents a set of approaches to explore an unknown asteroid with onboard calculations, and to land on its surface area selected in an optimal fashion. 
    The main motivation is to avoid the extended period of mapping phase or preliminary ground observations that are commonly required in spacecraft mission around asteroids. 
    First, range measurements from spacecraft to the surface are used to incrementally correct an initial shape estimate according to the Bayesian framework. 
    Then, an optimal guidance scheme is proposed to control the vantage point of the range sensor to construct a complete 3D model of the asteroid shape. 
    This shape model is then used in a nonlinear controller to track a desired trajectory about the asteroid.
    Finally, a multi resolution approach is presented to construct a higher fidelity shape representation in a specified location while avoiding the inherent burdens of a uniformly high resolution mesh. 
    This approach enables for an accurate shape determination around a potential landing site.
    We demonstrate this approach using several radar shape models of asteroids and provide a full dynamical simulation about asteroid 4769 Castalia.
    \keywords{Asteroids \and Shape Construction \and Guidance \and Optimization}
\end{abstract}

@@

% dynamics are difficult around asteroids
The dynamical environment around asteroids is strongly perturbed and challenging for analysis and mission operations~\cite{scheeres2012}.
Due to their low mass, which in turn causes a low gravitational attraction, asteroids may have irregular shapes.
Furthermore, asteroids may also have a chaotic spin state due to the absorption and emittance of solar radiation~\cite{rubincam2000}.
As a result, approaches utilizing an inverse square gravitational model do not capture the  true dynamical environment.
In addition, the vast majority of asteroids are difficult to track and characterize using ground based sensors.
Due to their small size, frequently with a maximum radius less than \SI{1}{\kilo\meter}, and low albedo, the reflected energy of these asteroids is insufficient for reliable detection or tracking.
Therefore, the dynamical model of the asteroid is relatively coarse prior to in situ measurements from a dedicated spacecraft.
As a result, any spacecraft mission to an asteroid must include the ability to update the dynamical model given in situ measurements and remain robust to unmodelled forces.

Another key dynamical consideration is the coupling between rotational and translational states around the asteroid.
The coupling is induced due to the different gravitational forces experienced on various portions of the spacecraft. 
The effect of the gravitational coupling is related to the ratio of the spacecraft size and orbital radius~\cite{hughes2004}.
For operations around asteroids, the ratio is relatively large which causes a much larger coupling between the translational and rotational states.
In~\cite{elmasri2005} and \cite{sanyal2004a}, the coupling of an elastic dumbbell spacecraft in orbit about a central body is considered, assuming that the central body is spherically symmetric.
Furthermore, the spacecraft model is assumed to remain in a planar orbit.
As a result, these developments are not directly applicable to motion about an asteroid, which experiences highly non-Keplerian motion.
In~\cite{misra2015b}, the effect of coupled motion is investigated for long term trajectories around asteroids.
However, the analysis only considered a second order spherical harmonic gravitational potential model. 
Therefore, these results are only valid when far from the asteroid surface and will diverge when used within the Brillouin sphere.

@@

% on arrival spacecraft spend long periods mapping, and depending on mission this might be unallocable
On approach to an asteroid, spacecraft navigation and guidance is primarily based on ground measurements.
After arrival, space will generally spend upwards of several months in a mapping mission phase~\cite{williams2018,kubota2003,cole1998}.
For example, the recent OSIRIS-REx mission used global imagery to create digital terrain maps over the span of approximately seven months~\cite{williams2018}.
During this period, spacecraft sensors, such as on board optical telescopes or Light radio Detection and Ranging (LIDAR), are used to characterize the asteroid.
The resulting imagery and range data is transmitted to the ground and the resulting asteroid shape and motion is estimated. 
During this mapping phase the spacecraft must remain in a quiescent state devoted entirely to mapping the surface.
Depending on the mission type, this long period of mapping is crucial to the mission, such as sample collection~\cite{williams2018}. 
For example, the OSIRIS-REx mission spent approximately seven months in mapping phase where digitial terrain maps and a shape model were constructed in support of main sample collection mission.
However, other missions, such as asteroid mitigation, may be severely limited by the time and ground resources required to generate a surface shape.
Furthermore, the long distances involved necessitate on-board autonomy to enable to spacecraft to operate without ground communications.
Similarly, during landing the spacecraft will require the ability to sense and model the surface topography in order to safely land in an unknown environment.
Onboard shape reconstruction techniques enable autonomous operations which allow for a greater range of spacecraft operations.
Mitigation of this ground based surface modeling will greatly expand the range of missions possible.

% TODO Reduce some of these paragraphs since they're wordy
% we want to generate the shape in real time and then use this shape for updated and better control
This paper presents three techniques to explore an asteroid without need for an extended period of mapping phase or ground based measurements. 
First, we develop a method to compute the surface shape of an asteroid from onboard range measurements.
Assuming that a meshed model of an initial estimate, such as an ellipsoidal shape, is available, the radius of any vertex in the measured area is adjusted according to the measurements. 
This is completed in a stochastic fashion where the degree of confidence in the current shape is compared with the level of measurement noise. 
Our approach is able to operate in real time and incrementally update the shape model of an asteroid as new range measurements are collected.
This approach allows for the shape to be continually updated as range measurements are used to locally modify the shape estimate.
This technique is verified by numerical examples of shape construction for asteroids Geographos and Golevka.

@@

The optimal guidance and multi resolution mapping schemes are illustrated by dynamical simulation of landing on the asteroid Castalia. 
During this simulation, this updated shape model is then used in a nonlinear controller to track a desired state trajectory for the dynamics of a rigid body spacecraft.
The dynamics are developed on the nonlinear manifold of rigid body motions, namely the special Euclidean group.
This formulation is based on an intrinsic geometric description of the motion and accurately captures the coupling between orbit and attitude dynamics.

@@

In this paper, we consider the motion of a dumbbell model of spacecraft around an asteroid.
The dumbbell model captures the important interactions of the coupling between orbital and attitude dynamics.
The dumbbell is defined by two spherical masses of radius \( r_1, r_2 \in \R \) with masses \( m_1, m_2 \in  \R\).
The masses are separated by a massless rod of length \( l \) and attached to the centers of each mass.
\Cref{fig:dumbbell_sc} shows the model and associated parameters.
\begin{figure}[htbp]
    \centering
    \input{dumbbell.tikz}
    \caption{Dumbbell Model of Rigid Spacecraft\label{fig:dumbbell_sc}}
\end{figure}
The dumbbell model is in some sense the simplest approximation of an extended rigid body.
This model captures the important dynamics of an extended rigid body with a minmum of additional complexity and is prevelant the literature~\cite{kulumani2017b,lee2006a,lee2007a,elmasri2005}.
In addition, the developments of the dumbbell model can be trivially extended to include a larger number of discrete masses to better approximate a more complicated spacecraft model.

Without loss of generality, we define body fixed frames for both the spacecraft and asteroid, which are aligned with the principle axes of each body and originate at their respective center of mass. 
The spacecraft body fixed frame is centered at the center of mass of the vehicle.
The \( b_1 \)  axis is aligned with the connecting rod and directed along the axis of symmetry.
The \( b_2, b_3 \) axes span the plane orthogonal to the axis of symmetry of the dumbbell.
The distance from the center of mass to each spherical mass is defined as
\begin{align}\label{eq:dumbbell_mass_distances}
    l_1 &= \frac{m_2}{m_1 + m_2} l, \\
    l_2 &= l - l_1.
\end{align}
The asteroid is modeled as a constant density polyhedron with constant, and known, spin about the axis of its maximum moment of inertia. 
The axes of the body fixed frame for the asteroid are denoted by $\{f_1,f_2,f_3\}$. 
We also define the inertial frame, whose axes are denoted by $\{e_1,e_2,e_3\}$.

@@

\subsection{Spacecraft Dynamical Model}

@@

Here, we summarize the key equations required for implementation.
The polyhedron potential is defined as 
\begin{align}
    U(\vc{r}) &= \frac{1}{2} G \sigma \sum_{e \in \text{edges}} \vc{r}_e \cdot \vc{E}_e \cdot \vc{r}_e \cdot L_e - \frac{1}{2}G \sigma \sum_{f \in \text{faces}} \vc{r}_f \cdot \vc{F}_f \cdot \vc{r}_f \cdot \omega_f,
\end{align}
and \( \vc{r}_e\) and \(\vc{r}_f \) are the vectors from the spacecraft to any point on the respective edge or face, \( G\in\Re\) is the universal gravitational constant, and \( \sigma\in\Re \) is the constant density of the asteroid.
We define the attraction, gravity gradient matrix, and Laplacian as
\begin{align}
    \nabla U ( \vc{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vc{E}_e \cdot \vc{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \vc{r}_f \cdot \omega_f \in \R^{3 \times 1} , \label{eq:attraction}\\
    \nabla \nabla U ( \vc{r} ) &= G \sigma \sum_{e \in \text{edges}} \vc{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \omega_f \in \R^{3 \times 3}, \label{eq:gradient_matrix}\\
    \nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f \in \R^1 .\label{eq:laplacian}
\end{align}
The Laplacian, defined in~\cref{eq:laplacian}, gives a simple method to determine if the spacecraft has collided with the body~\cite{werner1996}. 
The sum \( \sum \omega_f \) vanishes when outside the body and equals \( -4 \pi \) inside.
These equations are utilized to compute the gravitational potential of  the presented dumbbell spacecraft model.

@@

We assume the spacecraft contains a range sensor, such as LIDAR, that allows for the accurate measurement of the relative distance between the spacecraft and asteroid~\cite{zuber1997,zuber2000}.
Recent missions such as OSIRIS-REx have featured a scanning LIDAR system which allows for much higher measurement rates compared to previous missions~\cite{lauretta2017,daly2017}.
In these mission, ranging measurements are combined with imagery to derive a shape model using a ground based method~\cite{gaskell2008}.
In constrast to previous missions, the scanning LIDAR system on OSIRIS-REx enables for high surface sample rates in spite of the low ground-track velocities~\cite{daly2017}.
This type of sensor measures the round-trip time for a pulse of energy to leave the spacecraft, reflect off the surface, and return to a collector on board.
Given the time total time of flight, the distance can be accurately computed using \( d = \frac{\Delta TOF }{2} \) where \( c = \SI{2.998e8}{\meter\per\second}\) is the constant speed of light.
Assuming accurate knowledge of the pointing direction of the spacecraft, in the form of the rotation matrix \( R \in \R^{3 \times 3 } \), we can compute a direction from the spacecraft to the measurement location on the surface.
The output of this sensor is a vector, \( \vc{d}_i\in\Re^3 \), defined in the spacecraft fixed frame which gives the direction to a measurement point on the surface. 
Using the state of the asteroid, we can transform this measurement to the asteroid fixed frame using the simple transformation
\begin{align*}
    \vc{p}_i = R_A^T (x+R \vc{d}_i), 
\end{align*}
which is the vector from the origin of the asteroid fixed frame to the surface point of the measurement.

@@

Our algorithm applies a probabilistic framework to radially modify each vertex \( \vc{v}_i \in \R^3\) of the shape estimate based on measurement \( \vc{p}_i \in \R^3 \). 
In other words, each vertex from the initial estimate of shape is either stretched or shrunk along its radial direction, without any rotation. 
We assume that the initial intertial state of the vehicle is available. 
This data is typically available from a ground based orbit determination scheme which enabled the original orbital rendezvous. 
The approach presented here is reasonable assuming that both the initial estimate and the actual shape of the body does not have any hole. 
This approach alleviates much of complexity of incorporating new vertices or surface triangulation common in surface reconstruction methods~\cite{berg2008}.
This implies that the total number of vertices of the shape model is fixed.
However, additional detail, in the form of additional vertices, is possible by using standard mesh subdivision algorithms~\cite{orourke1998}, which is discussed in the subsequent section for multi resolution mapping.

@@

From spherical trigonometry~\cite{gade2010}, the central angle between measurement \( \vc{p}_j \) and vertex \( \vc{v}_i \) of the shape estimate is given by
\begin{align}\label{eq:geodesic_distance}
    \Delta \sigma_{j,i} = \arctan \parenth{\frac{\norm{\vc{p}_j \times \vc{v}_i}}{\vc{p}_j \cdot \vc{v}_i }}.
\end{align}
The parameters \( c \) and \( \Delta \sigma_{max} \) are used to define the region of impact of each measurement. 
The angular constraint \( \Delta \sigma_{max} \) may be used to adjust the computational requirements of the shape refinement process.
Larger values of \( \Delta \sigma_{max} \) result in a greater computational demand as a given measurement ray is used to update a larger number of vertices.
\Cref{sec:landing_refinement} presents an approach to allow for a higher resolution mesh in certain regions to capture finer surface details without excessive resources. 
The scalar \( c \) may then be used to adjust the variance \( w_{j,i} \) as a function of the distance between measurement \( j \) and vertex \( i \).
These two parameters may be used to minimize steep gradients that may occur during the shape reconstruction process.

@@

The results in this section utilize a kinematics only model of the spacecraft instead of the full dynamical simulation. 
We ignore the dynamics of the asteroid and spacecraft and instead focus solely on the shape reconstruction, and the spacecraft is assumed to be able to arbitrarily move around the asteroid and collect measurements.
The spacecraft remains at a fixed orbital radius defined as twice the maximum semi-major axis of the asteroid.
The simulations begin with an triaxial ellipsoid mesh that is sized to match the semi-major axes of each asteroid.
With this estimate, measurements are made of the surface and used to reconstruct the true shape following the process in~\cref{sec:radius_update}.
LIDAR measurements are generated until the total uncertainty of the model
\begin{align*}
    \sum_i w_i,
\end{align*}
is sufficiently small.
This uncertainty metric is the summation of the uncertainty of each vertex of the shape estimate.
As a result, a small total uncertainty is used to indicate that additional measurements would not substantially modify the shape.
In addition, we compute the volume of the estimated shape and compare against the volume of the truth shape model~\cite{neese2004}.

@@

The control gains are chosen based on the desired closed-loop system response. 
A variety of techniques are available to choose these gains, but a simple linear analysis offers a straightforward and systematic approach to choosing suitable values. 
We use the control inputs defined in~\cref{eq:translational_control,eqn:nodist_control} and substitute them into the dynamical equations of motion in~\cref{eq:x_dot,eq:v_dot,eq:R_dot,eq:omega_dot}.

@@

We utilize radar shape models of asteroid \num{4769} Castalia and (\num{52760}) \num{1998} \(\text{ML}_{14}\)~\cite{neese2004}.
The examples demonstrate the full dynamical simulation of a rigid spacecraft with an autonomous closed loop control scheme to both reconstruct the asteroid shape.
The proposed optimal guidance scheme is executed in the outer-loop to generate the desired trajectory, which is followed by the nonlinear control scheme in the inner-loop. 
More specifically, the nonlinear controllers described previously are used to control both the translational and rotational states of the vehicle, where the control inputs are computed using the current shape estimate of the asteroid. 
As measurements are collected, the spacecraft autonomously updates its shape estimate and uses this estimate to compute the control inputs and desired future states.
As such, both the shape model and the gravitational potential available to the controller will be gradually refined. 
However, throughout the simulation, the actual spacecraft dynamics are computed with the gravity computed by the full shape model unknown to the controller.
These demonstrate the ability of a spacecraft to autonomously explore and maneuver around an initially poorly modeled asteroid while incrementally updating the shape model.

In this numerical example we have assumed the perfect knowledge of the spacecraft state. 
This is a strong assumption that would not exist in any realistic situation. 
However, the measurement model defined in the previous section may be extended to consider the uncertainty in both the state \( x \) and the range measurement \( d \) rather than combining them into a single random variable \( p \).
In addition, the dynamics defined by~\crefrange{eq:x_dot}{eq:omega_dot} do not consider any non-gravitational accelerations such as solar radiation pressure on the vehicle.
In general, these types of external forces will primarily serve to increase state uncertainty and cause large errors in the closed loop control system.
Adaptive control techniques have been demonstrated in the past to address both fixed and time varying disturbances of this kind~\cite{kulumani2017a}.
However, these topics are beyond the scope of this paper focusing on optimal guidance for Bayesian shape reconstruction, and relegated to future works.

The asteroids are assumed to constantly rotate about the \( \vc{f}_3 = \vc{e}_3\) axis according to the parameters given in~\cref{tab:dynamic_asteroids}.
Furthermore, the state of the asteroid, namely the rotation matrix \( R_A \), is assumed to be known based on ground measurements or previous data.
\begin{table}[htbp]
    \centering
    \begin{tabular}{lcc}
        \toprule
        Property & \num{4769} Castalia & (\num{52760}) \num{1998} \(\text{ML}_{14}\) \\
        \midrule
        Semi-major axes(\si{\kilo\meter}) & \( 0.8065 \times 0.4905 \times 0.413 \) & \( 1.1 \times 1.1 \times 1.1 \) \\
        Rotational Period (\si{\hour}) & \num{4.095} & \num{14.98} \\
        Density (\si{\gram\per\centi\meter^3}) & \num{2.1} & \num{2.1} \\
        Vertices & \num{2048}  & \num{8192} \\
        Faces & \num{4092} & \num{16320} \\
        \bottomrule
    \end{tabular}
    \caption{Asteroid properties for dynamical exploration~\label{tab:dynamic_asteroids}}
\end{table}
At the beginning of the simulation the spacecraft is assumed to lie on the inertial \( e_1 \) axis, i.e.\ \( x_0 = [ x_0,\, 0,\,  0 ]\, \si{\kilo\meter} \).
In addition, at the initial state the spacecraft is orientated such that the \( b_1 \) axis is aligned with the inertial \( e_2 \) axis.
In other words the initial orientation is given by \( R_0 = \exp(\frac{\pi}{2} \hat{e}_3)\).
The shape reconstruction phase of the simulation is performed over \SI{15000}{\second}, over which time the spacecraft will take LIDAR measurements of the surface at \SI{1}{\hertz}.
Once the total uncertainty has been reduced sufficiently the spacecraft maneuvers to a ``home'' position aligned with the \( f_1 \) axis of the asteroid.

@@

Utilizing these metrics, a landing site is chosen to minimize the surface cost given as
\begin{align}\label{eq:surface_cost}
    \min_{v_i} J_l(x, v_i) =  J_{\text{distance}}(x, v_i) - J_{\text{science}}(v_i),
\end{align}
subject to a hard inequality constraint requiring that the surface slope is less than a threshold, i.e., \( \phi \leq \phi_m \).
The surface cost described here is relatively simple and chosen to be solely a function of the current spacecraft position and shape model.
Additional scientific criteria such as surface material or location on the body, e.g. equatorial vs. polar regions, can be utilized instead of a random science value as presented here.
